---
title: "Lab_9_assignment:Rmarkdown scientific report"
author: "Ella Cavicchi"
date: "2026-03-26"
output: 
  html_document:
    toc: true
    toc_float: true
    theme: flatly
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# YOUR TO DO LIST :

### How you present and organize your markdown is flexible, get creative with it! This is what you should include:

-   Deduce and summarize aspects of its journey :

    -   Its start and end locations
    -   The latitudinal change (difference between max and min latitude)
    -   The total distance your shark traveled in km

-   Include the following markdown elements somewhere in your report :

    -   Headings
    -   Bold text
    -   Italicized text
    -   An embedded figure of a basking shark (find online)
    -   A table

-   **Knit your document and submit as an HMTL on brightspace**

# **Load and Filter**

```{r load-packages, include=FALSE}
packages <- c("tidyverse", "sf", "rnaturalearth", "rnaturalearthdata")

# Install packages not yet installed
installed_packages <- packages %in% installed.packages()

if (any(installed_packages == FALSE)) {
  install.packages(packages[!installed_packages])
}

# Packages loading
invisible(lapply(packages, library, character.only = TRUE))
```

## Define Coasts and Ocean

```{r load-spatial-data, echo=FALSE}
# Country polygons for the Americas
coast <- ne_countries(scale = "medium", returnclass = "sf")

# Ocean polygon
ocean <- ne_download(scale = "medium", 
                     type = "ocean", 
                     category = "physical",
                     returnclass = "sf")
```

## Load Data

```{r load-shark-data}
# Download shark tracking data directly, ERDDAP servers often go down for maintenance, looks like that's currently the case 
#sharks <- read.csv("https://erddap.bco-dmo.org/erddap/tabledap/bcodmo_dataset_476315.csv", 
                   #skip = 1) 

#We'll use the csv
sharks <- read.csv("~/Desktop/BIOL3872/Lab 9/Data/basking_sharks.csv")
sharks <- sharks[-1,]

# Clean column names
colnames(sharks) <- c("ptt", "tuid", "year", "month", "day", "longitude", "latitude", 
                      "date", "type", "depth_min", "temp_depth_min", "depth_max", 
                      "depth_range", "tsst")

# Convert date column and add month name
sharks <- sharks %>%
  mutate(
    across(c(ptt, tuid, year, month, day, longitude, latitude, 
             depth_min, temp_depth_min, depth_max, depth_range, tsst), as.numeric),
    date = as.Date(date, format = "%m/%d/%Y"),
    month_name = month.abb[month]
  )


# Convert to spatial object
sharks_sf <- st_as_sf(sharks, 
                      coords = c("longitude", "latitude"),
                      crs = 4326)
```

## Filter Data for Shark of Interest

```{r your-shark}

my_shark <- 52562

# Filter data
my_data <- sharks %>%
  filter(ptt == my_shark) %>%
  arrange(date)

my_sf <- sharks_sf %>%
  filter(ptt == my_shark) %>%
  arrange(date)
```

## Spatial Polygons
```{r bbox}
bbox <- st_bbox(my_sf)
xlim <- c(bbox["xmin"] - 3, bbox["xmax"] + 3)
ylim <- c(bbox["ymin"] - 3, bbox["ymax"] + 3)
```

# **Path of Shark 52562**

```{r create plot, echo=FALSE, fig.width=7, fig.height=9}
# Edit the settings of this chunk to render appropriately, what do you need to add to {r}?

# Build the map layer by layer
my_data %>% mutate(month_name = factor(month_name, levels = month.abb))%>%
  ggplot() +
  # Layer 1: Ocean background
  geom_sf(data = ocean, fill=NA) +
  
  # Layer 2: Land polygons
  geom_sf(data=coast, fill="gray60")+ 
  
  # Layer 3: Track line connecting points
  geom_path(aes(x=longitude, y=latitude))+
  
  # Layer 4: Points colored by month
  geom_point(aes(x=longitude, y=latitude, color=month_name)) +
  
  # Layer 5: Start point (green triangle)
  geom_point(data=my_data[which.min(my_data$date),], aes(x=longitude, y=latitude), shape=17, size=4.5, color= "green", alpha=0.75)+
  
  # Layer 6: End point (red square)
  geom_point(data=my_data[which.max(my_data$date),], aes(x=longitude, y=latitude), shape=15, size=4.5, color="red", alpha=0.75) +
  
  # Set map extent
  coord_sf(xlim = xlim, ylim = ylim, expand = FALSE) +
  scale_color_manual(
  name = "Month",
  values = c("Jan" = "#313695", "Feb" = "#4575b4", "Mar" = "#74add1",
             "Apr" = "#abd9e9", "May" = "#fee090", "Jun" = "#fdae61",
             "Jul" = "#fd8c1c", "Aug" = "#e7400d", "Sep" = "#d73027",
             "Oct" = "#f46d43", "Nov" = "#a50026", "Dec" = "#922f80")
  )+
  labs(
    title = paste("Migration of Basking Shark", 52562),
    x = "Longitude", y = "Latitude"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(size = 16, face = "bold"),
    legend.position = "right"
  )
```

```{r session-info}
sessionInfo()
```
